---
title: "Epidemiology of Co-Colonized Methicillin Resistant Staphylococcus aureus Associated ABSSSI"
author: "Richard Evans"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    code_folding: hide
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

```{r knitr_init, echo = FALSE, cache = FALSE, warning = FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print = "75")
opts_chunk$set(echo = TRUE,
	             cache = TRUE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE)
opts_knit$set(width = 85)
```

# Introduction

R language Rmarkdown document to accompany *Epidemiology of Co-Colonized Methicillin Resistant Staphylococcus aureus Associated Acute Bacterial Skin and Skin Structure Infection*.

Cleaning of study data began in SAS v9.2 and project migrated to R shortly thereafter. Several sections of this analysis code did not contribute to the manuscript, they are left in the document to demonstrate data exploration and possible methods of analysis.

# Libraries

```{r}
library(haven)       # import data - sas7bdat
library(tidyverse)   # suite of data cleaning, processing, management functions
library(magrittr)    # add() function, and `%<>%`
library(visdat)      # visualize whole data frames
library(tableone)    # CreateTableOne package
library(knitr)       # Rmarkdown tables
library(kableExtra)  # make Rmarkdown tables pretty
```

# Source Data

```{r, message = FALSE, warning = FALSE}
setwd("S:/MartinEpi/Projects/Co-Infx/project/data")

files <- list.files(pattern = "*.sas7bdat")
file.names <- gsub(".sas7bdat", "", files)

for (i in 1:length(files)) {
  assign(file.names[i], read_sas(files[i]))
}
```

# Process Data

```{r}
# dput(colnames(main_fixed))
dates_to_convert <- c("adm_dat", "cult_dat")
numeric_to_factor <- c("male", "obese", "eth", "race", "resid", "ltcf", "wound",
                       "neutro", "surg_yr", "abx_yr", "device", 
                       "diabetes", "diab_com", "amputat", 
                       "retino", "neuro", "nephro", "hba_4", "renal", "chr_wnd", "wnd_site", 
                       "wnd_pre", "steroid", "chemothp", "radiothp", "post_trn",
                       "ulcer", "ulc_stg1", "ulc_stg2", "ulc_stg3", 
                       "ulc_stg4", "cssti", "sepsis", "new_icu", "new_int", "new_surg", 
                       "new_proc", "new_cvc", "new_uc", "new_arf", "new_liv", "new_endo", 
                       "vaso", "abx_indx", "res_sus", "gram", "aerobe", "poly_infx", 
                       "res_org", "bacter", "iso_org", "gram_r", "aerobe_r", "post_stat", 
                       "fail_dth", "read_30", "infx_doc", "trt_fail", 
                       "fail_res1", "fail_res2", 
                       "fail_icu", "fail_org", "fail_reop", "fail_infx", 
                       "fail_tot", "eld", "ulc", "extremity", "dth_study", "cssti1", 
                       "cssti2", "cssti3", "cssti4", "year", "hosp", "black", "white", 
                       "other", "home", "ltacf", "oth_resid", "emia", "EMDTH", "kidney", 
                       "charl_cat", "cssti_lev", "vre",
                       "switch", "switch_30", "dth_30"
                       )

tmp_fn <- function(x) as.Date(x, origin = "1960-01-01")

main <- main_fixed %>%
  mutate(
    hba_4 = ifelse(hba_4 == '.' | hba_4 == '', NA, hba_4),
    ulc_stg2 = ifelse(ulc_stg2 == '', NA, ulc_stg2),
    ulc_stg3 = ifelse(ulc_stg3 == '', NA, ulc_stg3),
    ulc_stg4 = ifelse(ulc_stg4 == '', NA, ulc_stg4),
    fail_res2 = ifelse(fail_res2 == '', NA, fail_res2),
    fail_dat = as.Date(fail_dat, "%m/%d/%Y")
  ) %>%
  mutate_at(dates_to_convert, funs(tmp_fn)) %>%
  mutate_at(numeric_to_factor, funs(factor))
```

# Explore

## Table 1

### Continuous variables age and BMI
```{r}
print(CreateTableOne(vars = c("age", "bmi"),
                     strata = "poly_infx",
                     data = main),
      nonnormal = c("age", "bmi"))
```

### Categorical variables

```{r}
# dput(colnames(main))
cont_vars <- c("male","obese", "eth", "race", "resid", "ltcf", "wound",
              "neutro", "surg_yr", "abx_yr", "device",
              "diabetes", "diab_com", "amputat",
              "retino", "neuro", "nephro", "hba_4", "renal", "chr_wnd", 
              "wnd_site", "steroid", "chemothp", "radiothp", "post_trn", 
              "ulcer", "ulc_stg1", "ulc_stg2", "ulc_stg3", 
              "ulc_stg4", "cssti", "sepsis", "new_icu", "new_int", "new_surg", 
              "new_proc", "new_cvc", "new_uc", "new_arf", "new_liv", "new_endo", 
              "vaso", "abx_indx", "res_sus", "gram", "aerobe",
              "res_org",  "bacter", "iso_org", "gram_r", "aerobe_r", 
              "post_stat", "fail_dth", "read_30",
              "infx_doc", "trt_fail", "fail_res1", "fail_res2", 
              "fail_icu", "fail_org", "fail_reop", "fail_infx", 
              "eld", "ulc", "extremity", "dth_study", "cssti1", 
              "cssti2", "cssti3", "cssti4", "year", "hosp", "black", "white", 
              "other", "home", "ltacf", "oth_resid", "emia", "EMDTH", "kidney", 
              "charl_cat", "cssti_lev", "switch", "switch_30", "dth_30")

print(CreateTableOne(vars = cont_vars,
                     strata = "poly_infx",
                     data = main,
                     argsApprox = list(correct = FALSE)))
```

## Categorize ABSSSI site of infection

```{r}
# table(main$infx_site) %>% 
#   as.data.frame() %>%
#   head()

# main %>% 
#   select(infx_site, infx_site1) %>% 
#   filter(infx_site1 == '')

upper_extrem <- "antecubital|arm|axilla|elbow|eblow|forearm|armpit|wrist"
lower_extrem <- "lower extremity|ankle|leg|knee|calf|heel|heal|malleolus|malleouls|patela|tibia|tiba|lower extremeties"
buttock <- "glute|buttock|butttock|coccyx|sacrum|sacral|scral|gloeal|tailbone"
hand <- "finger|hand|thumb"
foot <- "toe|foot|feet|hallux"
abdomen <- "abdomen|abdominal|abominal|peg|suprapubic|j tube"
upper_body <- "breast|chest|shoulder|shcoulder|soulder|sternal|deltoid"
back_sides <- "epidural|back|neck|psoas|cervical|lumbar|spine|spinal|hip|ischial|ischium|flank|side"
blood <- "blood|picc|line sepsis"
head <- "face|facial|chin|head|eye|brow|cheek|ear|parotid|temporal|nasal|nose|scalp|buccal|mandibula"
groin_genitals <- "groin|inguinal|thigh|pelvic|pelvis|vulva|vulvar|penile|penis|scrotal|scrotum|vagina|vaginal"
rectal <- "rectal|peri-rectal|rectal|perineum|perineal"
respiratory <- "sputum|respiratory|tracheal|tracheostomy|lungs|lung"

grepl_case <- function(pattern, x) {
  return(grepl(pattern, x, ignore.case = TRUE, perl = TRUE))
}

main <- main %>%
  mutate(
    infx_site = gsub("\\s+", " ", infx_site),
    site = ifelse(grepl_case(upper_extrem, infx_site), 'Upper Extremity', ''),
    site = ifelse(grepl_case(lower_extrem, infx_site), 'Lower Extremity', site),
    site = ifelse(grepl("LLE|LE|RLE", infx_site), "Lower Extremity", site),
    site = ifelse(grepl_case(buttock, infx_site), 'Buttock', site),
    site = ifelse(grepl_case(hand, infx_site), 'Hand', site),
    site = ifelse(grepl_case(foot, infx_site), 'Foot', site),
    site = ifelse(grepl_case(abdomen, infx_site), 'Abdomen', site),
    site = ifelse(grepl_case(upper_body, infx_site), 'Upper Body', site),
    site = ifelse(grepl_case(back_sides, infx_site), 'Back and Sides', site),
    site = ifelse(grepl_case(blood, infx_site), 'Blood', site),
    site = ifelse(grepl_case(head, infx_site), 'Head', site),
    site = ifelse(grepl_case(groin_genitals, infx_site), 'Groin and Genitals', site),
    site = ifelse(grepl_case(rectal, infx_site), 'Rectal', site),
    site = ifelse(grepl_case(respiratory, infx_site), 'Respiratory', site),
    infx_site1 = gsub("\\s+", " ", infx_site1),
    site1 = ifelse(grepl_case(upper_extrem, infx_site1), 'Upper Extremity', ''),
    site1 = ifelse(grepl_case(lower_extrem, infx_site1), 'Lower Extremity', site1),
    site1 = ifelse(grepl("LLE|LE|RLE", infx_site1), "Lower Extremity", site1),
    site1 = ifelse(grepl_case(buttock, infx_site1), 'Buttock', site1),
    site1 = ifelse(grepl_case(hand, infx_site1), 'Hand', site1),
    site1 = ifelse(grepl_case(foot, infx_site1), 'Foot', site1),
    site1 = ifelse(grepl_case(abdomen, infx_site1), 'Abdomen', site1),
    site1 = ifelse(grepl_case(upper_body, infx_site1), 'Upper Body', site1),
    site1 = ifelse(grepl_case(back_sides, infx_site1), 'Back and Sides', site1),
    site1 = ifelse(grepl_case(blood, infx_site1), 'Blood', site1),
    site1 = ifelse(grepl_case(head, infx_site1), 'Head', site1),
    site1 = ifelse(grepl_case(groin_genitals, infx_site1), 'Groin and Genitals', site1),
    site1 = ifelse(grepl_case(rectal, infx_site1), 'Rectal', site1),
    site1 = ifelse(grepl_case(respiratory, infx_site1), 'Respiratory', site1),
    site1 = ifelse(site1 == '', 'Unknown', site1),
    
    site1 = ifelse((site1 == '' & site == 'Buttock'), 'Buttock', site1),
    site1 = ifelse((site1 == '' & site == 'Blood'), 'Blood', site1),
    site1 = ifelse((site1 == '' & site == 'Foot'), 'Foot', site1)
    )
```

Good enough for now

```{r}
levs <- c("Head", "Upper Body", "Upper Extremity", "Hand", "Abdomen",
          "Back and Sides", "Buttock", "Rectal", "Groin and Genitals",
          "Lower Extremity", "Foot")

sites_orgs <- main %>%
  filter(poly_infx == 1) %>%
  select(id, site1, abau:citro, unidentified) %>%
  mutate(site1 = factor(site1, levels = levs))

totals <- data.frame(
  organism = names(sites_orgs[, -c(1, 2)]),
  sum = sapply(sites_orgs[, -c(1, 2)], sum)
  ) %>%
  mutate(perc = round((sum / sum(sum)) * 100, 1))

totals
```

## Categorize into aerobic vs. non-aerobic, gram (-) vs. (+), and fungal and mixture

```{r}
anaerobes <- c("b_fragilis", "bacteroid", "clost", "pepto")
gram_neg <- c("aeru", "ecoli", "abau", "mirab", "kpneu", "eaero", "ecloa",
              "proteus", "providencia", "morgan", "cfreu", "koxy", "afaecalis",
              "prevotella", "serratia", "malt", "vulgar", "eikenella", "hinflu",
              "mcatarr")
gram_pos <- c("coryne", "faecalis", "strep", "sagal", "staph", "sviridan",
              "faecium", "entero", "aureus", "lactobac", "spyogen", "sangin",
              "aturi", "micro", "mavium")
fungal <- c("albicans", "candida", "yeast", "fusarium")

indicate_orgs <- function(df, str) {
  ifelse(rowSums(df[, str]) >= 1, 1, 0)
}

main$anaerobic <- indicate_orgs(main, anaerobes)
main$gram_neg <- indicate_orgs(main, gram_neg)
main$gram_pos <- indicate_orgs(main, gram_pos)
main$fungal <- indicate_orgs(main, fungal)

types <- c("anaerobic", "gram_neg", "gram_pos", "fungal")
main <- main %>%
  mutate(
    infx_type = ifelse(anaerobic == 1, 'Anaerobic', ''),
    infx_type = ifelse(gram_neg == 1,  'Gram Negative', infx_type),
    infx_type = ifelse(gram_pos == 1,  'Gram Positive', infx_type),
    infx_type = ifelse(fungal == 1, 'Fungal', infx_type),
    infx_type = ifelse(anaerobic == 1 & gram_neg == 1, 
                       'Anaerobic, Gram Negative',
                       infx_type),
    infx_type = ifelse(anaerobic == 1 & gram_pos == 1,
                       'Anaerobic, Gram Positive',
                       infx_type),
    infx_type = ifelse(anaerobic == 1 & fungal == 1,
                       'Anaerobic, Fungal',
                       infx_type),
    infx_type = ifelse(gram_neg == 1 & gram_pos == 1,
                       'Gram Negative, Gram Positive',
                       infx_type),
    infx_type = ifelse(gram_neg == 1 & fungal == 1,
                       'Gram Negative, Fungal',
                       infx_type),
    infx_type = ifelse(gram_pos == 1 & fungal == 1,
                       'Gram Positive, Fungal',
                       infx_type),
    infx_type = ifelse(anaerobic == 1 & gram_neg == 1 & gram_pos == 1 & fungal == 1,
                       'Anaerobic, Gram Negative, Gram Positive, Fungal',
                       infx_type)
  )
```

```{r}
sapply(main[, types], table)

tab <- main %>%
  filter(poly_infx == 1 & infx_type != '') %>%
  select(infx_type, site1) %>%
  group_by(site1, infx_type) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(percent = count / sum(count) * 100)

tab
```

How many organisms in each polymicrobial infection? Evidently, and intuitively, frequency of organisms in a wound increases bioburden. Can test this with outcomes against total count of organisms in polymicrobial MRSA ABSSSI.

```{r}
main <- main %>%
  mutate(
    org_sum = rowSums(.[c(116:159, 161)]),
    org_sum = ifelse(poly_infx == 0, NA, org_sum)
  )

psych::describe(main$org_sum)
```

Fix VRE

```{r}
main <- main %>%
  mutate(
    vre = ifelse(grepl("VRE", resist), 1, 0),
    tmp = as.character(resist),
    tmp = tolower(tmp),
    tmp = gsub(" ", "", tmp),
    vre_faecalis = ifelse(grepl("vre.faecalis", tmp), 1, 0),
    vre_faecium = ifelse(grepl("vre.faecium", tmp), 1, 0)
  ) %>%
  select(-tmp)
```

```{r}
with(main, table(vre_faecalis, faecalis))
with(main, table(vre_faecium, faecium))
```

## By ABSSSI Site

### Overall abundance per site

```{r}
data.frame(table(main$site1))
```

By co-colonizing site

```{r}
sites <- main[main$poly_infx == 1, "site1"]
data.frame(table(sites))
```

### Most abundant Co-colonizing organisms by ABSSSI site

```{r}
co_colony.df <- main %>%
  filter(poly_infx == 1) %>%
  select(id, site1, abau:citro, unidentified, infx_type, cssti, diabetes)

co_colony.df %>%
  select(-id, -infx_type, -unidentified, -coryne, -cssti, -diabetes) %>%
  reshape2::melt(id.vars = "site1") %>%
  plyr::ddply(plyr::.(site1, variable),
              summarise,
              sum = sum(value)) %>%
  group_by(site1) %>%
  mutate(max = max(sum)) %>%
  filter(sum == max)
```

## Heatmaps

Define heatMap function

```{r, warning = FALSE, message = FALSE}
# import table
setwd("S:/MartinEpi/Projects/Co-Infx/project/documents/")
orgs <- readxl::read_excel("organism categories.xlsx", sheet = "type")
# head(orgs)

site_levels <- c("Head", "Upper Body", "Upper Extremity", "Hand",
                 "Abdomen", "Back and Sides",
                 "Buttock", "Rectal", "Groin and Genitals",
                 "Lower Extremity", "Foot")

# function to generate heatmaps for our specific needs
# requires ggplot2, magrittr
#' @param df data.frame object holding variables x, y, and fill
#' @param x vector of character or factor values for the x axis
#' @param y vector of character or factor values for the y axis
#' @param fill vector of numeric values for heat intensities
#' @param xlab label for x-axis
#' @param barheight height of the guide legend bar
#' @param barwidth width of the guide legend bar
#' @param legend_top position guide legend at top of heatmap
#' @param angle angle of x-axis ticks
#' @param yremove if TRUE, remove all y axis items

heatMap <- function(df, x, y, fill, xlab,
                    barheight = NULL, barwidth = NULL,
                    legend_top = FALSE,
                    angle = TRUE,
                    yremove = FALSE) {
  heat <- df %>%
    ggplot(aes_string(x = x, y = y, fill = fill)) %>%
    add(geom_tile()) %>%
    add(scale_fill_gradient(low = "navyblue",
                            high = "darkorange1")) %>%
    add(theme_minimal(14)) %>%
    add(scale_y_discrete(limits = rev(levels(df$variable)))) %>%
    add(labs(
      x = xlab,
      y = "",
      fill = "Proportion"
    ))
  if (angle) 
    heat <- heat + theme(
      axis.text.x = element_text(angle = 45,
                                 hjust = 1,
                                 face = "bold")
    )
  if (legend_top) 
    heat <- heat + 
      theme(legend.position = "top") +
      guides(fill = guide_colorbar(barwidth = barwidth,
                                   barheight = barheight))
  else
    heat <- heat + guides(fill = guide_colorbar(barheight = barheight))
  if (yremove) 
    heat <- heat + theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  return(heat)
}
```

### All organisms by MRSA ABSSSI site

Need to combine some sites due to sparsity

```{r}
co_colony.comb <- co_colony.df %>%
  mutate(
    site1 = ifelse(site1 == 'Head', 'Upper Body', site1),
    site1 = ifelse(site1 == 'Hand', 'Upper Extremity', site1),
    site1 = ifelse(site1 == 'Rectal', 'Buttock', site1),
    site1 = ifelse(site1 == 'Foot', 'Lower Extremity', site1)
  )
```

```{r}
site.heat1 <- co_colony.comb %>%
  filter(!(site1 %in% c('Respiratory', 'Unknown', 'Blood'))) %>%
  select(-id, -infx_type, -unidentified, -cssti, -diabetes) %>%
  reshape2::melt(id.vars = "site1") %>%
  mutate(
    site1 = factor(site1,
                   levels = site_levels,
                   labels = site_levels)
  ) %>%
  plyr::ddply(plyr::.(site1, variable),
              summarise,
              sum = sum(value)) %>%
  group_by(site1) %>%
  mutate(
    prop = (sum / sum(sum)) * 100,
    variable = as.character(variable),
    variable = factor(variable,
                      levels = orgs$varname,
                      labels = orgs$species)
    ) %>%
  ungroup()

site.heatmap1 <- heatMap(site.heat1, barheight = 31.75,
                         x = 'site1', y = 'variable', fill = 'prop',
                         xlab = "\n MRSA ABSSSI site")
site.heatmap1

# setwd("S:/MartinEpi/Projects/Co-Infx/project/documents/")
# tiff("site_heatmap1_all_comb.tiff",
#      width = 6, height = 8, units = "in", res = 300)
# site.heatmap1
# dev.off()
```

### MRSA ABSSSI site by organism importance

```{r}
important <- orgs %>%
  filter(!(varname %in% c('providencia', 'afaecalis', 'prevotella', 'serratia', 
                          'eikenella', 'mcatarr', 'citro', 'coryne', 'lactobac',
                          'aturi', 'micro',' mavium', 'fusarium', 'yeast',
                          'hinflu', 'mavium', 'vulgar')))

site.heat2 <- co_colony.df %>%
  filter(!(site1 %in% c('Respiratory', 'Unknown', 'Blood'))) %>%
  select(-id, -infx_type, -unidentified, -cssti, -diabetes) %>%
  reshape2::melt(id.vars = "site1") %>%
  mutate(
    site1 = factor(site1,
                   levels = site_levels,
                   labels = site_levels)
  ) %>%
  filter(variable %in% important$varname) %>%
  plyr::ddply(plyr::.(site1, variable),
              summarise,
              sum = sum(value)) %>%
  group_by(site1) %>%
  mutate(
    prop = (sum / sum(sum)) * 100,
    variable = as.character(variable),
    variable = factor(variable,
                      levels = important$varname,
                      labels = important$species)
    ) %>% ungroup()

site.heatmap2 <- heatMap(site.heat2, barheight = 31.75,
                         x = "site1", y = "variable", fill = "prop",
                         xlab = "\n MRSA ABSSSI site")
site.heatmap2

# setwd("S:/MartinEpi/Projects/Co-Infx/project/documents/")
# tiff("site_heatmap2_important_comb.tiff",
#      width = 6, height = 8, units = "in", res = 300)
# site.heatmap2
# dev.off()
```

### All organisms by ABSSSI/cSSTI class

```{r}
cssti.heat1 <- co_colony.df %>%
  mutate(cssti = factor(cssti, 
                        levels = paste0(seq(1:4)),
                        labels = paste0(rep("class ", 4), 1:4))) %>%
  filter(!(site1 %in% c('Respiratory', 'Unknown', 'Blood'))) %>%
  select(-id, -infx_type, -unidentified, -site1, -diabetes) %>%
  reshape2::melt(id.vars = "cssti") %>%
  plyr::ddply(plyr::.(cssti, variable),
              summarise,
              sum = sum(value)) %>%
  group_by(cssti) %>%
  mutate(
    prop = (sum / sum(sum)) * 100,
    variable = as.character(variable),
    variable = factor(variable,
                      levels = orgs$varname,
                      labels = orgs$species)
    ) %>% ungroup()

cssti.heatmap1 <- heatMap(cssti.heat1, barheight = 35.6,
                          x = "cssti", y = "variable", fill = "prop",
                          xlab = "\n ABSSSI class",
                          angle = FALSE)

cssti.heatmap1

# setwd("S:/MartinEpi/Projects/Co-Infx/project/documents/")
# tiff("cssti_heatmap1_all_comb.tiff",
#      width = 6, height = 8, units = "in", res = 300)
# cssti.heatmap1
# dev.off()
```

### Important organisms by ABSSSI/cSSTI class

```{r}
cssti.heat2 <- co_colony.df %>%
  mutate(cssti = factor(cssti, 
                        levels = paste0(seq(1:4)),
                        labels = paste0(rep("class ", 4), 1:4))) %>%
  filter(!(site1 %in% c('Respiratory', 'Unknown', 'Blood'))) %>%
  select(-id, -infx_type, -unidentified, -site1, -diabetes) %>%
  reshape2::melt(id.vars = "cssti") %>%
  filter(variable %in% important$varname) %>%
  plyr::ddply(plyr::.(cssti, variable),
              summarise,
              sum = sum(value)) %>%
  group_by(cssti) %>%
  mutate(
    prop = (sum / sum(sum)) * 100,
    variable = as.character(variable),
    variable = factor(variable,
                      levels = important$varname,
                      labels = important$species)
    ) %>% ungroup()

cssti.heatmap2 <- heatMap(cssti.heat2, barheight = 35.6,
                          x = "cssti", y = "variable", fill = "prop",
                          xlab = "\n ABSSSI class",
                          angle = FALSE)

cssti.heatmap2

# setwd("S:/MartinEpi/Projects/Co-Infx/project/documents/")
# tiff("cssti_heatmap2_important.tiff",
#      width = 6, height = 8, units = "in", res = 300)
# cssti.heatmap2
# dev.off()
```

```{r, echo = FALSE, eval = FALSE}
cssti.heat3 <- cssti.heat1 %>%
  rename(site_or_cssti = cssti) %>%
  mutate(
    site_or_cssti = as.character(site_or_cssti),
    variable = as.character(variable),
    type = 'MRSA ABSSSI Class'
  ) %>%
  select(-sum)

site.heat3 <- site.heat1 %>%
  rename(site_or_cssti = site1) %>%
  mutate(
    site_or_cssti = as.character(site_or_cssti),
    variable = as.character(variable),
    type = "MRSA ABSSSI Site"
  ) %>%
  select(-sum)

levs_and_labs <- union(levels(site.heat1$site1), 
                       levels(cssti.heat1$cssti))

heat3 <- rbind(site.heat3, cssti.heat3) %>%
  mutate(
    variable = factor(variable,
                      levels = orgs$species,
                      labels = orgs$species),
    site_or_cssti = factor(site_or_cssti,
                           levels = levs_and_labs,
                           labels = levs_and_labs),
    type = factor(type)
  ) %>%
  ggplot(aes(x = site_or_cssti, y = variable, fill = prop)) %>%
  add(geom_tile()) %>%
  add(scale_fill_gradient(low = "navyblue",
                          high = "darkorange1")) %>%
  add(facet_wrap(~type, scales = "free_x")) %>%
  add(scale_y_discrete(limits = levels(site.heat1$variable))) %>%
  add(theme_minimal(base_size = 18)) %>%
  add(guides(fill = guide_colorbar(barheight = 48.5))) %>%
  add(theme(axis.text.x = element_text(angle = 330, hjust = 0))) %>%
  add(labs(
    x = "",
    y = "",
    fill = "Proportion"
    ))
heat3

# tiff("csstiXsite_hmap_all_comb.tiff",
#      width = 12, height = 12, units = "in", res = 700, compression = "lzw")
# heat3
# dev.off()
```

```{r, echo = FALSE, eval = FALSE}
cssti.heat4 <- cssti.heat2 %>%
  rename(site_or_cssti = cssti) %>%
  mutate(
    site_or_cssti = as.character(site_or_cssti),
    variable = as.character(variable),
    type = 'MRSA ABSSSI Class'
  ) %>%
  select(-sum)

site.heat4 <- site.heat2 %>%
  rename(site_or_cssti = site1) %>%
  mutate(
    site_or_cssti = as.character(site_or_cssti),
    variable = as.character(variable),
    type = "MRSA ABSSSI Site"
  ) %>%
  select(-sum)

levs_and_labs <- union(levels(site.heat2$site1), 
                       levels(cssti.heat2$cssti))

heat4 <- rbind(site.heat4, cssti.heat4) %>%
  mutate(
    variable = factor(variable,
                      levels = important$species,
                      labels = important$species),
    site_or_cssti = factor(site_or_cssti,
                           levels = levs_and_labs,
                           labels = levs_and_labs),
    type = factor(type)
  ) %>%
  ggplot(aes(x = site_or_cssti, y = variable, fill = prop)) %>%
  add(geom_tile()) %>%
  add(scale_fill_gradient(low = "navyblue",
                          high = "darkorange1")) %>%
  add(facet_wrap(~type, scales = "free_x")) %>%
  add(scale_y_discrete(limits = rev(levels(site.heat2$variable)))) %>%
  add(theme_minimal(18)) %>%
  add(guides(fill = guide_colorbar(barheight = 42))) %>%
  add(theme(axis.text.x = element_text(angle = 330, hjust = 0))) %>%
  add(labs(
    x = "",
    y = "",
    fill = "Proportion"
    ))

heat4

# tiff("csstiXsite_hmap_important.tiff",
#      width = 10, height = 10, units = "in", res = 700, compression = "lzw")
# heat4
# dev.off()
```

# Polymicrobial distribution analyses

multi-response permutation procedure (MRPP)

## Distribution differences by cSSTI/ABSSSI classification

Removing 'rare' species

```{r}
csstiXorg <- co_colony.df %>%
  select(cssti, aeru, ecoli, abau, mirab, kpneu, ecloa, proteus, coryne, strep,
         sagal, staph, faecium, faecalis, albicans, candida) %>%
  mutate(
    sumVar = rowSums(.[2:15]),
    cssti = factor(cssti,
                   levels = paste0(1:4),
                   labels = paste0(rep("Class ", 4), 1:4))) %>%
  filter(sumVar > 0) %>%
  select(-sumVar) %>%
  na.omit() %>%
  as.data.frame()
```

Performing the MRPP

```{r}
library(vegan)
set.seed(1233)
cssti.mrpp <- mrpp(csstiXorg[-1], csstiXorg$cssti, permutations = 1000)
cssti.mrpp
```

The MRPP provides a test of whether there is a significant difference between two or more groups of sampling units. Similar to components analysis, we are examining the differences in communities between ABSSSI/cSSTI classes. The p-value for the $\delta$ statistic from this procedure indicates a non-significant difference in community (p = `r cssti.mrpp$pvalue`). $\delta$ itself is the group mean difference calculated from each of the permuted datasets (N = 999).

Mean distance/mean within and between block dissimilarities. This calculates a matrix of mean within-cluster dissimilarities (the diagonal) and between-cluster dissimilarities (off-diagonal entries), and an attribute n of grouping counts.

```{r}
cssti.md <- meandist(vegdist(csstiXorg[-1]), csstiXorg$cssti)
summary(cssti.md)
```

```{r}
par(mfrow = c(1, 2))
with(cssti.mrpp, {
  fig.dist <- hist(boot.deltas, xlim = range(c(delta, boot.deltas)),
                   main = "Test of Differences Among Groups",
                   xlab = "Bootstrapped Delta")
  abline(v = delta);
  text(delta, 2*mean(fig.dist$counts), adj = -0.5,
       expression(bold(delta)), cex = 1.5)
})
plot(cssti.md, kind = "histogram", 
     main = "Within-group and between-group dissimilarities")
```

In the right-hand plot above, the horizontal line is drawn at the level of mean between-cluster dissimilarity and vertical lines connect within-cluster dissimilarities to this line.

## Distribution differences by diabetic status

Same treatment for diabetic patients (vs. non-diabetics)

```{r}
diabXorg <- co_colony.df %>%
  select(diabetes, aeru, ecoli, abau, mirab, 
         kpneu, ecloa, coryne, strep, sagal, 
         staph, faecium, faecalis, albicans, 
         candida) %>%
  mutate(
    sumVar = rowSums(.[2:15]),
    diabetes = factor(diabetes,
                      levels = c(0, 1),
                      labels = c("Non-Diabetic", "Diabetic"))
    ) %>%
  filter(sumVar > 0) %>%
  select(-sumVar) %>%
  na.omit() %>%
  as.data.frame()
```

```{r}
set.seed(1233)
diab.mrpp <- mrpp(diabXorg[-1], diabXorg$diabetes, permutations = 1000)
diab.mrpp
```

Mean distance by diabetic grouping

```{r}
diab.md <- meandist(vegdist(diabXorg[-1]), diabXorg$diabetes)
summary(diab.md)
```

```{r}
plot(diab.md, kind = "dendrogram", 
     main = "MRSA ABSSSI Community Dendrogram - Diabetic Status")
text(2, 0.865, "MRPP p-val < 0.01")
```

A ggplot2 version of this dendrogram

```{r, echo = FALSE, eval = TRUE}
# true minimal theme for plotting ggplot objects
theme_trueMinimal <- function(base_size = 16, base_family = "") {
    theme_bw(base_size = base_size, base_family = base_family) %+replace% 
        theme(axis.ticks = element_blank(), 
              legend.background = element_blank(), 
              legend.key = element_blank(), 
              panel.background = element_blank(), 
              strip.background = element_blank(), 
              panel.border = element_blank(),
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(),
              plot.background = element_blank(), 
              axis.line = element_line(), 
              axis.line.y = element_blank(),
              complete = TRUE)
}
```

```{r}
dendro <- data.frame(
  x = c(0.4, 1), 
  y = c(0.8636266, 0.8701293),
  label = c("Non-Diabetic", "Diabetic")
  ) %>% 
  ggplot(aes(x = x, y = y, label = label)) %>% 
  add(geom_segment(aes(x = 0.4, xend = 1, y = 0.8805288, yend = 0.8805288))) %>%
  add(geom_segment(aes(x = 0.4, xend = 0.4, y = 0.8805288, yend = 0.8636266))) %>%
  add(geom_segment(aes(x = 1, xend = 1, y = 0.8805288, yend = 0.8701293))) %>%
  add(geom_text(vjust = 1)) %>%
  add(annotate("text", x = 1, y = 0.865, label = "MRPP p-val < 0.01")) %>%
  add(expand_limits(x = c(-0.3, 1.5))) %>%
  add(theme_minimal(16)) %>%
  add(labs(
    x = "", 
    y = "Mean Distance\n",
    title = ""
    )) %>%
  add(theme(
    axis.text.x = element_blank(),
    axis.line.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ))
dendro
```

```{r}
par(mfrow = c(1, 2))
with(diab.mrpp, {
  fig.dist <- hist(boot.deltas, xlim = range(c(delta, boot.deltas)),
                   main = "Test of Differences Among Groups",
                   xlab = "Bootstrapped Delta")
  abline(v = delta);
  text(delta, 2 * mean(fig.dist$counts), adj = -0.5,
       expression(bold(delta)), cex = 1.5)
})
plot(diab.md, kind = "dendrogram", 
     main = "Within-group and between-group dissimilarities")
```

Diabetics may have different distributions of organisms in their MRSA ABSSSIs (p = `r diab.mrpp$Pvalue`).

Let's look at a heatmap of co-colonizing organisms faceted by diabetic status to see if we can identify differences

```{r}
diab.heat1 <- co_colony.df %>%
  mutate(diabetes = factor(diabetes, 
                      levels = c(0, 1),
                      labels = c("Non-diabetic", "Diabetic"))) %>%
  filter(!(site1 %in% c('Respiratory', 'Unknown', 'Blood'))) %>%
  select(-id, -infx_type, -unidentified, -cssti, -site1) %>%
  reshape2::melt(id.vars = "diabetes") %>%
  plyr::ddply(plyr::.(diabetes, variable),
              summarise,
              sum = sum(value)) %>%
  group_by(diabetes) %>%
  mutate(
    prop = (sum / sum(sum)) * 100,
    variable = as.character(variable),
    variable = factor(variable,
                      levels = orgs$varname,
                      labels = orgs$species)
    ) %>%
  ungroup() %>%
  ggplot(aes(x = diabetes, y = variable)) %>%
  add(geom_tile(aes(fill = prop))) %>%
  add(scale_fill_gradient(low = "navyblue",
                          high = "darkorange1")) %>%
  add(theme_minimal(18)) %>%
  add(scale_y_discrete(limits = rev(levels(site.heat1$variable)))) %>%
  add(guides(fill = guide_colorbar(barheight = 30))) %>%
  add(labs(
    x = "",
    y = "",
    fill = "%"
    ))

diab.heat1

# ggsave("diabetes_hmap_all.tiff", diab.heat1,
#        device = "tiff",
#        width = 8, height = 8, units = "in",
#        dpi = 700, compression = "lzw")
```

```{r, echo = FALSE, eval = FALSE}
# g <- cowplot::plot_grid(dendro, diab.heat1,
#                         labels = c("A", "B"),
#                         ncol = 1,
#                         rel_heights = c(1.1, 3.0))
# 
# ggsave("heat_dendro.tiff", g,
#        device = "tiff",
#        width = 8, height = 12, units = "in", dpi = 700,
#        compression = "lzw")
```

```{r, echo = FALSE, eval = FALSE}
# setwd("S:/MartinEpi/Projects/Co-Infx/project/documents")
# tiff("community_analyses.tiff",
#      width = 12, height = 6, units = "in", res = 300)
# par(mfrow = c(1, 2))
# plot(cssti.md, kind = "dendrogram",
#      main = "Community Difference Dendrogram - cSSTI class")
# text(1.2, 0.845, "MRPP p-val = 0.07")
# 
# plot(diab.md, kind = "dendrogram",
#      main = "Community Differences Dendrogram - Diabetic status")
# text(2, 0.863, "MRPP p-val = <0.01")
# dev.off()
```

```{r}
factors <- orgs$species[c(5:9, 11, 26, 28:30, 32, 27, 41, 42)]

diab.heat2 <- diabXorg %>%
  reshape2::melt(id.vars = "diabetes") %>%
  plyr::ddply(plyr::.(diabetes, variable),
              summarise,
              sum = sum(value)) %>%
  group_by(diabetes) %>%
  mutate(
    prop = (sum / sum(sum)) * 100,
    variable = factor(variable,
                      labels = factors)
    ) %>%
  ungroup() %>%
  ggplot(aes(x = diabetes, y = variable)) %>%
  add(geom_tile(aes(fill = prop))) %>%
  add(scale_fill_gradient(low = "navyblue",
                          high = "darkorange1")) %>%
  add(theme_minimal(18)) %>%
  add(scale_y_discrete(limits = rev(factors))) %>%
  add(guides(fill = guide_colorbar(barheight = 30))) %>%
  add(labs(
    x = "",
    y = "",
    fill = "%"
    ))

diab.heat2
```

```{r, echo = FALSE, eval = FALSE}
# marge <- theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
# g <- cowplot::plot_grid(dendro + marge, 
#                         diab.heat2 + marge,
#                         labels = c("A", "B"),
#                         ncol = 1,
#                         rel_heights = c(1.1, 3.0))
# 
# ggsave("heat_dendro.tiff", g,
#        device = "tiff",
#        width = 8, height = 12, units = "in", dpi = 700,
#        compression = "lzw")
```

### Indicator species analysis

indicator species analysis using the methods of Legendre & Dufrene. The goal of indicator analysis is to determine which species can be used as indicators of certain site groups. The method calculates an index that measures the association between a species and a site group. Specifically, the index is calculated between the species for diabetics and non-diabetics and looks for the group corresponding to the highest association value, the statistical significance of this relationship is tested using a permutation test.

```{r}
library(indicspecies)
indic <- multipatt(diabXorg[-1], diabXorg$diabetes, control = how(nperm = 999))
summary(indic)
```

*A. baumanii* is strongly and significantly associated with non-diabetic polymicrobial MRSA ABSSSI while *Staphylococcus spp.* are significantly associated with diabetic polymicrobial MRSA ABSSSI.

indicator value components

```{r}
summary(indic, indvalcomp = TRUE)
```

The indicator value index is the product of two components, A and B:

* A: p(surveyed site belongs to target site | species has been found)
    + equivalent to *specificity* or *positive predictive value*
* B: p(finding the species in sites belonging to the side group)
    + equivalent to *sensitivity* or *fidelity*

Here, *A. baumannii* is a good indicator for non-diabetic polymicrobial MRSA ABSSSI because it occurs in 80% of non-diabetic wounds and only 18% in diabetic wounds.

Pearson's $\phi$ coefficient of association for binary vectors can be used similar to the above:

```{r}
indic_cor <- multipatt(diabXorg[-1], 
                       diabXorg$diabetes,
                       func = "r.g", 
                       control = how(nperm = 999))
summary(indic_cor)
```

*E. faecalis*, when examining the heatmap above, does appear to be more abundant in diabetic wounds, so, this analysis does make some sense.

The differences between indicator analyses and correlation analyses are important: **correlation indices** are helpful for determining the preferences of species among groups (diabetics and non-diabetics) while **indicator value indices** are used for assessing the predictive values of species as indicators of those groups (diabetics vs. non-diabetics).

# Risk factor predictive modelling

```{r}
main_ml <- main %>%
  select(male, poly_infx, eld, obese, black, ltacf, diabetes,
         kidney, device, charl_cat, chr_wnd, surg_yr, abx_yr, 
         cssti_lev, cult_days, steroid, ulc)
```

Impute missing data

```{r}
pMiss <- function(x) (sum(is.na(x)) / length(x)) * 100
apply(main_ml, 2, pMiss)
```

```{r, message = FALSE, warning = FALSE}
library(mice)
main_impute <- mice(main_ml, m = 5, maxit = 50, meth = "pmm", printFlag = FALSE)
main_imputed <- complete(main_impute, 1)
```

testing and training

```{r}
partition <- caret::createDataPartition(main_imputed$poly_infx,
                                        p = 0.7,
                                        list = FALSE)
main_train <- main_imputed[partition, ]
main_test <- main_imputed[-partition, ]
dim(main_train); dim(main_test)

# some models require all variables to be numeric (no factors)
main_train_num <- main_train %>%
  mutate_all(funs(as.numeric)) %>%
  mutate_at(vars(male:cssti_lev, steroid:ulc), funs(. - 1))

main_test_num <- main_test %>%
  mutate_all(funs(as.numeric)) %>%
  mutate_at(vars(male:cssti_lev, steroid:ulc), funs(. - 1))
```

## Fast and Frugal Trees

```{r}
library(FFTrees)
coinfx.fft <- FFTrees(poly_infx ~ .,
                      data = main_train_num,
                      data.test = main_test_num,
                      main = "Co-Colonized MRSA ABSSSI",
                      decision.labels = c("Mono-Microbial", "Polymicrobial"))
```

```{r}
coinfx.fft
summary(coinfx.fft)
```

```{r, fig.height = 12, fig.width = 10}
plot(coinfx.fft,
     data = "test",
     cue.cex = 1.5,
     cue.labels = c("Ulcer", "LTCF", "Charlson \n 4+", "Device"),
     decision.cex = 1.5,
     show.header = FALSE,
     label.tree = "Decision Tree")
```

```{r, echo = FALSE, eval = FALSE}
# tiff("coinfx_fft.tiff",
#      height = 10, width = 14, units = "in",
#      res = 800, compression = "lzw")
# plot(coinfx.fft,
#      data = "test",
#      cue.cex = 1.5,
#      cue.labels = c("Ulcer", "LTCF", "Charlson \n 4+", "Hx Chronic \n Wounds"),
#      decision.cex = 1.5,
#      show.header = FALSE,
#      label.tree = "Fast and Frugal Tree")
# dev.off()
```

```{r}
plot(coinfx.fft, what = "cues")
```

```{r, echo = TRUE, eval = FALSE}
coinfx.fff <- FFForest(poly_infx ~ .,
                       data = rbind(main_train_num, main_test_num),
                       ntree = 100,
                       train.p = 0.5,
                       goal = "acc")
```

```{r, echo = FALSE, eval = TRUE}
setwd("S:/MartinEpi/Projects/Co-Infx/project/documents/")
# saveRDS(coinfx.fff, "coinfx_fff.rds")
coinfx.fff <- readRDS("coinfx_fff.rds")
coinfx.fff
```

```{r}
plot(coinfx.fff, node.cex.lim = c(8, 16), line.cex.lim = c(0.8, 10))
```

```{r}
# tiff("forest_of_FFTs.tiff", height = 6, width = 14, units = "in",
#      res = 800, compression = "lzw")
# plot(coinfx.fff,
#      node.cex.lim = c(8, 16),
#      line.cex.lim = c(0.8, 10))
# dev.off()
```

The above Fast and Frugal tree and forest suggests LTCF residence, Presentation of Ulcers, and a Charlson Score $\ge$ 4 as the 'biggest' predictors of a polymicrobial MRSA ABSSSI (included in 75% of the generated fast and frugal trees). A cSSTI level of 4+ and the existence of a medical device were so-so predictors. The network plot indicates that LTCF, Ulcers, Charlson's Score greater than 4, cSSTI class of 3 or 4, and implanted medical devices co-occured within each of the 100 FFTs.

```{r, fig.height = 6, fig.width = 10}
labels <- c("ABX past yr.",
            "Charlson 4+",
            "cSSTI 3+",
            "Med Device",
            "Diabetes",
            "Age 65+",
            "LTCF",
            "Surgery past yr.",
            "Ulcer")

colorset <- RColorBrewer::brewer.pal(9, "Set1")

tabletheme <- gridExtra::ttheme_minimal(
  core = list(fg_params = list(cex = 2.0)),
  colhead = list(fg_params = list(cex = 2.0)),
  rowhead = list(fg_params = list(cex = 2.0))
)

confusion <- t(coinfx.fff$forest.stats$train[c(6:9, 11, 13)])
confusion <- data.frame(Metric = round(confusion, 2))

FFFimportance <- coinfx.fff$frequencies %>%
  data.frame() %>%
  mutate(Var1 = factor(Var1, labels = labels)) %>%
  ggplot(aes(x = reorder(Var1, -Freq), 
             y = Freq, 
             fill = Var1)) %>%
  add(geom_col(color = "black", alpha = 0.7)) %>%
  add(geom_text(aes(label = Freq), size = 7, vjust = -1)) %>%
  add(theme_trueMinimal(18)) %>%
  add(theme(legend.position = "none")) %>%
  add(scale_fill_brewer(palette = "Set1", direction = -1)) %>%
  add(labs(
    x = "Variable",
    y = "Importance (%)"
  )) %>%
  add(expand_limits(y = c(0, 110))) %>%
  add(annotation_custom(gridExtra::tableGrob(confusion,
                                             theme = tabletheme),
                        xmin = 8.5, xmax = 8.5, ymin = 60, ymax = 100))
FFFimportance
```

```{r, echo = FALSE, eval = FALSE}
# setwd("S:/MartinEpi/Projects/Co-Infx/project/documents/")
# tiff("FFFimportance.tiff", height = 6, width = 14, units = "in",
#      res = 800, compression = "lzw")
# FFFimportance
# dev.off()
```

## Logistic regression

```{r}
library(h2o)
h2o.init()
predictors <- colnames(main_train)[-2]
response <- "poly_infx"
df.train <- as.h2o(main_train)
df.test <- as.h2o(main_test)

coinfx.lr <- h2o.glm(family = "binomial", x = predictors, y = response,
                     training_frame = df.train, lambda = 0,
                     compute_p_values = TRUE)

coinfx.lr@model$coefficients_table

h2o.std_coef_plot(coinfx.lr)
perf <- h2o.performance(coinfx.lr, df.test)
h2o.confusionMatrix(perf)
```

Accuracy is `r  round(((59 + 80) / (80 + 50 + 16 + 59)) * 100, 2)`%

The error rate is `r round((1 - (59 + 80) / (80 + 50 + 16 + 59)) * 100, 2)`%

## Gradient Boosting and Random Forest

2 - model ensemble: Gradient Boosting and Random Forest

```{r}
# train and cross-validate GBM
coinfx.gbm <- h2o.gbm(x = predictors, y = response,
                      training_frame = df.train,
                      distribution = "bernoulli",
                      ntrees = 10,
                      max_depth = 3,
                      min_rows = 2,
                      learn_rate = 0.2,
                      nfolds = 5,
                      fold_assignment = "Modulo",
                      keep_cross_validation_predictions = TRUE)

perf <- h2o.performance(coinfx.gbm, df.test)
h2o.confusionMatrix(perf)
h2o.varimp_plot(coinfx.gbm)
```

Accuracy is `r  round(((57 + 93) / (93 + 37 + 18 + 57)) * 100, 2)`%

The error rate is `r round((1 - (57 + 93) / (93 + 37 + 18 + 57)) * 100, 2)`%

```{r}
# train and cross-validate RF
coinfx.rf <- h2o.randomForest(x = predictors, y = response,
                              training_frame = df.train,
                              ntrees = 100,
                              nfolds = 5,
                              fold_assignment = "Modulo",
                              keep_cross_validation_predictions = TRUE)

perf <- h2o.performance(coinfx.rf, df.test)
h2o.confusionMatrix(perf)
h2o.varimp_plot(coinfx.rf)
```

Accuracy is `r  round(((78 + 65) / (78 + 52 + 10 + 65)) * 100, 2)`%

The error rate is `r round((1 - (78 + 65) / (78 + 52 + 10 + 65)) * 100, 2)`%

Worse than GBM

## LASSO/Elastic-Net

```{r}
detach('package:FFTrees')
detach('package:purrr')
detach('package:h2o')
library(glmnet)

y <- as.integer(main_train_num[, "poly_infx"])
x <- data.matrix(main_train_num[, c(1, 3:15)])

lasso1 <- glmnet(x = x,
                 y = y,
                 family = "binomial",
                 alpha = 0.5)

par(mfrow = c(1, 3))
plot(lasso1, xvar = "dev", label = TRUE)
plot(lasso1, xvar = "lambda", label = TRUE)
plot(lasso1, xvar = "norm", label = TRUE)
```

Copied from the *glmnet* vignette:

Prediction is a little different for logistic from Gaussian, mainly in the option type. "link" and "response" are never equivalent and "class" is only available for logistic regression. In summary, 

    * "link" gives the linear predictors
    * "response" gives the fitted probabilities
    * "class" produces the class label corresponding to the maximum probability.
    * "coefficients" computes the coefficients at values of s
    * "nonzero" retuns a list of the indices of the nonzero coefficients for each value of s.

For "binomial" models, results ("link", "response", "coefficients", "nonzero") are returned only for the class corresponding to the second level of the factor response.

Making predictions at class labels of $\lambda \in \left\{0.05, 0.01\right\}$ where $\lambda$ is the regularization parameter, that which controls the overall strength of the penalty. The penalty determines coefficient 'shrinkage' to zero and serves as our selection mechanism.

```{r}
pred1 <- predict(lasso1,
                 newx = data.matrix(main_test_num[, c(1, 3:15)]),
                 type = "class",
                 s = c(0.05, 0.01))

pred.tab <- table(main_test_num[,"poly_infx"], as.numeric(pred1[, 1]))
pred.tab
```

accuracy with $\lambda = 0.05$ `r round(sum(diag(pred.tab)) / sum(pred.tab) * 100, 2)`%.

```{r}
pred.tab <- table(main_test_num[,"poly_infx"], as.numeric(pred1[, 2]))
pred.tab
```

accuracy with $\lambda = 0.01$ is `r round(sum(diag(pred.tab)) / sum(pred.tab) * 100, 2)`%.

Cross-Valdiation with 10-folds using misclassification error as the criterion:

```{r}
lasso.cv <- cv.glmnet(x = x, y = y, family = "binomial", type.measure = "class")
plot(lasso.cv)
```

```{r}
coef(lasso.cv, s = "lambda.1se")
pred2 <- predict(lasso.cv, 
                 newx = data.matrix(main_test_num[, c(1, 3:15)]),
                 type = "class",
                 s = "lambda.1se")
pred.tab <- table(main_test_num[,"poly_infx"], as.numeric(pred2[, 1]))
pred.tab
```

accuracy with $\lambda = 0.02$ is `r round(sum(diag(pred.tab)) / sum(pred.tab) * 100, 2)`%. This is not as good, however, prediction this this model is not really the primary goal, the goal is to select a sparse matrix of important cofficients.

Cross-Valdiation with 10-folds using AUC as the criterion (ROC curve):

```{r}
lasso.cv <- cv.glmnet(x = x, y = y, family = "binomial", type.measure = "auc")
plot(lasso.cv)
```

# Outcome modelling

Table 4 in Evans **et al.**

Univariate

```{r}
outcomes <- c("read_30", "new_icu", "new_int", 
              "new_surg", "new_cvc", "new_uc", "dth_30")
outcome.df <- main %>%
  mutate_at(vars(outcomes), funs(as.numeric)) %>%
  mutate_at(vars(outcomes), funs(. - 1)) %>%
  as.data.frame()

psych::describe(main$los)
par(mfrow = c(1, 2))
plot(density(main$los)); hist(main$los)
```

Trying negative binomial regression

```{r}
library(MASS)
los_nb <- glm.nb(los ~ poly_infx, data = main)
summary(los_nb)

test <- glm(los ~ poly_infx, family = "poisson", data = main)
pchisq(2 * (logLik(los_nb) - logLik(test)), df = 1, lower.tail = FALSE)

RR <- exp(cbind(estimate = coef(los_nb), confint(los_nb)))
RR
```

having a polymicrobial MRSA ABSSSI increases the expected log count of the number of days in the hospital by 0.74 vs.a mono-microbial MRSA ABSSSI.

As for fit, check the poisson model,

```{r}
boot::glm.diag.plots(test)
```

Against the negative binomial model,

```{r}
boot::glm.diag.plots(los_nb)
```

Much better fit.

```{r}
los_nb <- glm.nb(los ~ poly_infx + age + chr_wnd + charl_cat + cssti_lev,
                 data = main)
summary(los_nb)

test <- glm(los ~ poly_infx + age + chr_wnd + charl_cat + cssti_lev,
            family = "poisson",
            data = main)
pchisq(2 * (logLik(los_nb) - logLik(test)), df = 1, lower.tail = FALSE)
boot::glm.diag.plots(los_nb)

RR <- exp(cbind(estimate = coef(los_nb), confint(los_nb)))
RR

rm(los_nb, test, RR)
detach('package:MASS')
```

```{r, warning = FALSE, message = FALSE}
lapply(outcome.df[, outcomes],
       function(var) {
         mod <- glm(var ~ poly_infx, data = outcome.df, family = binomial)
         OR <- exp(mod$coefficients)
         CI <- exp(confint(mod))
         cbind(OR, CI)
       })
```

Table 5 in Evans **et al.**

Multivariate Outcomes

```{r, warning = FALSE, message = FALSE}
lapply(outcome.df[, outcomes],
       function(var) {
         mod <- glm(var ~ poly_infx + age + chr_wnd + charl_cat + cssti_lev,
                    data = outcome.df,
                    family = binomial)
         n <- length(mod$residuals)
         prop <- round((n / 688) * 100, 2)
         OR <- exp(mod$coefficients)
         CI <- exp(confint(mod))
         cbind(OR, CI, n, prop)
       })
```

Stratifying/adjusting this analysis by ABSSSI site

```{r}
# reduce sites to simpler categories
main.sites <- main %>%
  mutate(
    site = ifelse(site1 == 'Head', 'Upper Body', site1),
    site = ifelse(site == 'Hand', 'Upper Extremity', site),
    site = ifelse(site == 'Rectal', 'Buttock', site),
    site = ifelse(site == 'Foot', 'Lower Extremity', site),
    site = ifelse(site %in% c('Blood', 'Respiratory'), 'Unknown', site)
  ) %>%
  mutate_at(vars(outcomes), funs(as.numeric)) %>%
  mutate_at(vars(outcomes), funs(. - 1)) %>%
  filter(site != 'Unknown') %>%
  as.data.frame()

table(main.sites$site)
dim(main.sites)
```

```{r, warning = FALSE, message = FALSE}
lapply(main.sites[, outcomes],
       function(var) {
         mod <- glm(var ~ poly_infx + age + chr_wnd + charl_cat + cssti_lev + site,
                    data = main.sites,
                    family = binomial)
         n <- length(mod$residuals)
         prop <- round((n / 688) * 100, 2)
         OR <- exp(mod$coefficients)
         CI <- exp(confint(mod))
         cbind(OR, CI, n, prop)
       })
```

## Stratified analysis

Subsets analysis

```{r, warning = FALSE, message = FALSE}
site_cats <- unique(main.sites$site)
stratLogit <- function(var) {
  form <- as.formula(paste0(var, "~ poly_infx+age+chr_wnd+charl_cat+cssti_lev"))
  models <- vector("list", length(site_cats))
  for (i in 1:length(site_cats)) {
    name <- site_cats[i]
    mod <- glm(formula = form,
               data = main.sites,
               subset = (site == site_cats[i]),
               family = binomial)
    n <- length(mod$residuals)
    prop <- round((n / 688) * 100, 2)
    OR <- round(exp(mod$coefficients)[2], 2)
    LB <- round(as.numeric(exp(confint(mod))[2, 1]), 2)
    UB <- round(as.numeric(exp(confint(mod))[2, 2]), 2)
    models[[i]] <- cbind(OR, LB, UB, n, prop)
  }
  models <- do.call(rbind, models)
  row.names(models) <- site_cats
  return(models)
}

modelLists <- vector("list", length(outcomes))
for(i in 1:length(outcomes)) {
  modelLists[[i]] <- stratLogit(outcomes[i])
}
names(modelLists) <- outcomes

modelLists
```

Firth's logistic regression to handle sparsity/separation between predictor and outcome

```{r}
library(logistf)

stratLogistf <- function(var) {
  form <- as.formula(paste0(var, "~ poly_infx+age+chr_wnd+charl_cat+cssti_lev"))
  models <- vector("list", length(site_cats))
  for (i in 1:length(site_cats)) {
    name <- site_cats[i]
    mod <- logistf(formula = form,
                   data = main.sites[main.sites$site == site_cats[i], ])
    n <- mod$n
    prop <- round((n / 688) * 100, 2)
    OR <- round(exp(mod$coefficients)[2], 2)
    LB <- round(as.numeric(exp(confint(mod))[2, 1]), 2)
    UB <- round(as.numeric(exp(confint(mod))[2, 2]), 2)
    models[[i]] <- cbind(OR, LB, UB, n, prop)
  }
  models <- do.call(rbind, models)
  row.names(models) <- site_cats
  return(models)
}

modelLists <- vector("list", length(outcomes))
for(i in 1:length(outcomes)) {
  modelLists[[i]] <- stratLogistf(outcomes[i])
}
names(modelLists) <- outcomes

modelLists
```

## Severity of outcomes vs. polymicrobial frequency

```{r}
psych::describe(main$org_sum)
```

```{r, fig.height = 4, fig.width = 4}
hist(main$org_sum)
```

The median number of organisms in polymicrobial wounds was 2, ranging from 1 to 11.

How do the odds of our outcomes change as the number of organisms in the wound increase?

```{r}
models <- lapply(outcome.df[, outcomes],
                 function(var) {
                   mod <- glm(var ~ org_sum, 
                              data = outcome.df,
                              family = binomial,
                              subset = (poly_infx == 1))
                   OR <- exp(coefficients(mod)[2])
                   CI <- exp(confint(mod)[2, ])
                   c(colnames(var), OR, CI)
                   }
                 )
models
```

It really doesn't affect the occurrence of each individual outcome. What about the number of individual outcomes as an indicator of severity?

```{r}
temp <- outcome.df %>%
  mutate(sum_outcomes = rowSums(.[c(171, 194, 93:95, 97:98)])) %>%
  filter(!is.na(sum_outcomes))

test <- glm(sum_outcomes ~ org_sum, data = temp)
exp(coefficients(test))
exp(confint(test))
```

The answer is ... no. No, there is not any indication that, with more microbes inhabiting a polymicrobial MRSA ABSSSI, there will be more severe outcomes or more 'adverse' outcomes. Perhaps because all of these infections already have MRSA in them. Yeah, that's probably it.

# Misc. Functions

```{r}
# function to generate heatmaps for our specific needs
# requires ggplot2, magrittr
#' @param df data.frame object holding variables x, y, and fill
#' @param x vector of character or factor values for the x axis
#' @param y vector of character or factor values for the y axis
#' @param fill vector of numeric values for heat intensities
#' @param xlab label for x-axis
#' @param barheight height of the guide legend bar
#' @param barwidth width of the guide legend bar
#' @param legend_top position guide legend at top of heatmap
#' @param angle angle of x-axis ticks
#' @param yremove if TRUE, remove all y axis items

heatMap <- function(df, x, y, fill, xlab,
                    barheight = NULL, barwidth = NULL,
                    legend_top = FALSE,
                    angle = TRUE,
                    yremove = FALSE) {
  heat <- df %>%
    ggplot(aes_string(x = x, y = y, fill = fill)) %>%
    add(geom_tile()) %>%
    add(scale_fill_gradient(low = "navyblue",
                            high = "darkorange1")) %>%
    add(theme_minimal(14)) %>%
    add(scale_y_discrete(limits = rev(levels(df$variable)))) %>%
    add(labs(
      x = xlab,
      y = "",
      fill = "Proportion"
    ))
  if (angle) 
    heat <- heat + theme(
      axis.text.x = element_text(angle = 45,
                                 hjust = 1,
                                 face = "bold")
    )
  if (legend_top) 
    heat <- heat + 
      theme(legend.position = "top") +
      guides(fill = guide_colorbar(barwidth = barwidth,
                                   barheight = barheight))
  else
    heat <- heat + guides(fill = guide_colorbar(barheight = barheight))
  if (yremove) 
    heat <- heat + theme(
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  return(heat)
}

# true minimal theme for plotting ggplot objects
theme_trueMinimal <- function(base_size = 16, base_family = "") {
    theme_bw(base_size = base_size, base_family = base_family) %+replace% 
        theme(axis.ticks = element_blank(), 
              legend.background = element_blank(), 
              legend.key = element_blank(), 
              panel.background = element_blank(), 
              strip.background = element_blank(), 
              panel.border = element_blank(),
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(),
              plot.background = element_blank(), 
              axis.line = element_line(), 
              axis.line.y = element_blank(),
              complete = TRUE)
}
```

# Session Info

```{r}
sessionInfo()
```
